### 2023-01-14 rev1_cover_letter.tex
edis plus cover letter add

\documentclass[10pt,parskip=full,foldmarks=off,addrfield=off,backaddress=false,refline=dateleft,letterpaper]{scrlttr2}
\LoadLetterOption{kentucky}

\usepackage{xspace}

% Sender's information
\setkomavar{signature}{%
  \includegraphics[scale=0.3]{hand_sig.pdf}\\%
  Jeremy Van Cleve\\%
  Associate Professor\\%
  Department of Biology\\%
  University of Kentucky} % use PDF image of signature
\setkomavar{date}{\today}
\setkomavar{subject}{}

\usepackage[explicit]{titlesec}
\titleformat{\section}{\scshape \bfseries}{}{0ex}{#1}
\titlespacing{\section}{0pt}{1.0ex plus 0.25ex minus 0.25ex}{0.5ex plus 0.1ex minus 0.1ex}

\newenvironment{reviewerquote}{\begin{quote}\color{DarkBlue}\itshape}{\end{quote}}
\newenvironment{response}{}{}

\begin{document}

% Recipient's information
\begin{letter}

  \opening{Dear Phil Trans B}

  Please find enclosed my revision of the manuscript entitled ``Evolutionarily stable strategy analysis and its links to demography and genetics through invasion fitness''. The Reviewers made a number of helpful and probing comments, and I have attempted to address each one thoroughly below. Reviewer comments are quoted in blue italics and my responses immediately follow. I have also uploaded a ``tracked changes'' version of the PDF to enable easy reference for changes that have been made in response to the reviews.

  \closing{Best regards,}

\section{Editor comments}

\begin{reviewerquote}
  Reviewer 1 has made several useful suggestions for clarity and they should be straightforward to implement in the revision.

  While Reviewer 2 also finds the manuscript well written, they are more critical and make several comments which should be addressed and may require more extensive revision.
\end{reviewerquote}

I would like to thank the Editor and Reviewers for their very detailed in insightful comments. I have copied each Reviewer's comments below followed by a detailed response with changes I've made in response to the comments.

\section{Referee 1 comments}

\begin{reviewerquote}
  This paper reviews the history and current state of the field of ESS analysis. The author links ESS analysis to the concept of invasion fitness, and discusses the application of invasion fitness to kin selection and evolution in variable environments Although I learned some valuable and interesting things, and I enjoyed reading the first half of this paper, I also found it difficult to understand some of the latter parts of the paper. Since reviews usually aim to clarify and give a (relatively) simple overview of the literature, I think the paper would benefit from some rewrites and from more careful notation and explanation of notation.
\end{reviewerquote}

Thank you for the excellent feedback. I've endeavored to improve the latter parts of the paper and provide clearer explanations. Please see my responses below to the reviewer's specific comments.

\begin{reviewerquote}
Confusion/clarification/notational questions:
- You are using $x$ and $y$ to denote both the phenotypes, and their frequencies, and I’m not always sure which one you mean. E.g. up until line 138, x and y have been defined as denoting the phenotypes, so I assume that $w(x,y)$ in line 127 means that $x$ and $y$ are interacting, and it does not mean that w is a function of the frequencies of $x$ and $y$? It would help me a lot if you used different symbols for these two things, e.g. maybe $p_x$ and $p_y$ for the frequencies?
- Line 135: So now we have x denoting both the phenotype and the frequency of the phenotype, and then we have $x_{ij}$ denoting mating preference of genotype $A_i$ $A_j$. Is it possible to use a different symbol for $x_{ij}$?
\end{reviewerquote}

Thanks for mentioning this. I did start the section talking about two strategies, $x$ and $y$, and thus do mean as suggested above that $w(x,y)$  is a function of phenotypic values $x$ and $y$ and not of frequencies of those strategies in the population. However, I think what may be causing confusion is that then I moved to allowing individuals to choose strategies $x$ and $y$ probabilistically, so now mixed strategies are allowed and the strategy is technically the probability of selecting strategy $x$ (with complementary probability of selecting strategy $y$). Since the probabilities $x_{ij}$ are strategies, I'd like to keep the same notation $x$ for the probabilities.

However, a reworking of this section and some notation change I hope will make things simpler and clearer. I've started off assuming in section 2 that the strategies are probabilities of selecting behavior $C$ when individuals have two possible behaviors, $C$ and $D$. Now, the equation on line 138 (now eqn 2) is more clearly described as the mean strategy in the population or mean probability of selecting behavior $C$ and is no longer confusingly described as a frequency. I've added additional explanation in sections 2 and 3 to make the description of the phenotypes clearer.

\begin{reviewerquote}
- I’m not sure I understand what “the ESSs obtained by applying condition (1) to the payoff matrix $A$” means. Does it mean that we are considering some trait to evolve until the entries of the pay off matrix satisfies the ESS conditions, are x and y the evolving phenotypic trait? Does it means that at some frequencies of x and y, the entries of the pay off matrix satisfy the ESS condition (related to the previous question)?
\end{reviewerquote}

Yes, I agree that this is unclear. I have rewritten this paragraph to better describe both the population genetic model and the ESS analysis. The ESS analysis is simply to use the condition (1) to determine which of the two strategies is ESS or if there is a mixed strategy ESS. It should be clear now that the ESS value of the phenotype is the same value of the genotype frequencies that are stable in the population genetic model.

\begin{reviewerquote}
  - Line 184: Equation for mean fitness: what happened to $y$? Shouldn’t the mean fitness in the population also be a function of phenotype $y$?
\end{reviewerquote}

The equation is correct as written and the confusion here stems from the earlier issue the reviewer highlighted. In this scenario, the phenotype $x$ is the probability of selecting behavior $C$ (which I was calling ``$x$'' before confusingly) and thus the probability of selecting behavior $D$ is simply $1-x$. Thus, only one phenotypic variable, $x$, is needed to calculate the mean fitness. I have added additional clarification around this equation in addition to the earlier text clarifying the phenotypes.

\begin{reviewerquote}
  - Line 182: “the fitness function is linear”, linear in what/a linear function of what?
\end{reviewerquote}

I've had added that the fitness $w(x,y)$ is a linear function of its arguments, which are the phenotypes of the two interacting individuals. Linearity here is reasonable given the phenotype is a probability of selecting among two pure strategies.

\begin{reviewerquote}
  - Equation 3: $u_j$ is not yet defined I think? $u_{xy}$ has been defined as an entry of the pay off matrix, but this $u$ only comes with one index?
\end{reviewerquote}

\begin{reviewerquote}
- Line 212-213: “under the right assumptions”, and earlier in that paragraph “given the general genetic and demographic assumptions above”. I read through the above again to try and find what these assumptions are but had trouble locating them, except for maybe “the fitness function is linear” and “mutant allele is rare”? Can you repeat/list the assumptions necessary for Result 1 to hold?
- I guess $u_k(y,x)$ on line 305 is an element in the vector $u(y,x)$ from line 292?
- What are the dimensions of the matrix $A(y,x)$ in line 287? I’ve never encountered vectors used as matrix indices before (as in $a_{k’k}$, usually matrix indices are scalars), and I’m not sure how to interpret the notation? If $A$ contains all possible $k$ and $k’$ vectors, its dimensions must quickly get very large?
- Really enjoyed the section on variable environments.
Minor comments:
- Line 53: I’m not sure what it means for a method to be orthogonal to issues
Typos:
- Line 87: exits -> exists
- Line 97: in that in can in
- Line 155 an genetic -> a genetic
- Line 347 “One of the important features of Result 1 is that shows how”
\end{reviewerquote}

\section{Referee 2 comments}

\begin{reviewerquote}

This manuscript attempts to review the 50-year history of the ESS concept. In view of the more than 50,000 “ESS articles” published since 1973, this is a highly ambitious task. Although I agree with some of the take-home messages of the manuscript, I am not entirely convinced that Van Cleve has succeeded in his mission.

From the start, the ESS history was multi-faceted. The 1973 paper of Maynard Smith and Price was based on an earlier manuscript of Price (which I could inspect when visiting Maynard Smith many years ago). Here, the main novel idea was not the ESS concept but rather “strategy thinking”, adopted from game theory (strategy = recipe for condition-dependent behaviour). Price argued that a “mouse” strategy (nowadays called Dove) could be rapidly outcompeted by a more aggressive strategy, but that this would not be the case for a ”retaliator” strategy, which, in the absence of aggressive individuals, leads to precisely the same behaviour as the “mouse” strategy. From this, Price drew the (for him) most important conclusion that the “organisation of behaviour” (that is, the “strategy” underlying behaviour) is crucial for understanding behavioural evolution. In contrast, Maynard Smith had one main goal: to publish his ESS concept before Hamilton could work out his ideas on “unbeatable strategies” in more detail.

I mention all this because Von Cleve starts talking about evolutionary “game theory”, while game aspects (like strategy thinking) play a minor role in his review. He is not really interested in behavioural interactions, but in generalizing the ESS concept to more complicated settings. At first, I found this quite confusing, as, from the historical perspective, it is crucial that the ESS concept was first developed in a game context. In fact, Maynard Smith’s derivation of the ESS conditions (1a) and (1b) only makes sense for two-person matrix games, as in this case the fitness function w(x,y) is linear in both components. Various authors later “corrected” this “mistake” and derived more general ESS concepts for non-linear fitness functions. It is a pity that all this is not mentioned by Van Cleve. Indeed, non-linearities arise, for example, in “games between relatives”. How to deal with this (and, hence, how to capture kin selection in the ESS formalism) led to a vivid exchange of arguments between Grafen and Maynard Smith in the late 1970s and the early 1980s. Grafen pointed out that his approach (corresponding to an “actor-modulated approach”) leads to quite different outcomes than Maynard Smith’s inclusive fitness approach. As kin selection is a central theme of Van Cleve’s manuscript, I would have expected to hear about these early debates.

Anyway, it is now quite obvious that Maynard Smith would have designed the ESS conditions quite differently if he had considered the possibility of a non-linear fitness function. It soon became clear that such functions can have quite different evolutionary properties, like “Garden of Eden” scenarios where an ESS is not attainable, or “evolutionary branching” scenarios where the population converges to an evolutionarily unstable strategy. Later work on this resulted in Adaptive Dynamics theory, which is not mentioned at all in Van Cleve’s manuscript, although the invasion fitness concept of Adaptive Dynamics is more sophisticated than the variant propagated by Van Cleve. Interestingly, from the Adaptive Dynamics perspective, the second ESS condition of Maynard Smith is not a condition for “evolutionary stability” at all, but rather a condition for “convergence stability”. This has been noticed before (in writing) by various colleagues, including Eshel, Lessard, Metz, and Dieckmann. It would have been very interesting to hear Van Cleve’s position regarding this reinterpretation of “ESS history”.

The middle part of Van Cleve’s review (Sections 3 and 4) contains some interesting observations, but the results are often stated in a relatively sloppy manner (i.e., without mentioning the conditions under which they hold). In fact, the references referred to in the review all make quite limiting assumptions, such as a linear fitness function (in both components) or a system with only two phenotypes. Take, for example, the discussion of internal versus internal stability in Section 4, two very interesting concepts indeed. These concepts are mainly useful in systems with many different strategies. In contrast, the central Result 1 of that Section only applies to systems with two pure strategies (or phenotypes), and it does not hold anymore in three-strategy games like Rock-Scissors-Paper (this is, for example, explicitly mentioned in the key reference 103). Obviously, this limitation of Result 1 should have been mentioned. In fact, various counter-examples to both parts of Result 1 were constructed in the evolutionary branch of economic game theory (most notably by Friedman and Nachbar). Like my colleagues from economics, I also spent several years on attempts to extend Result 1 to more than two phenotypes. Like them, I had to conclude that long-term evolution does not “save” the ESS concept when more than two strategies are around.

But also the case of only two phenotypes is less clear than presented by Van Cleve. A good example can be found in the excellent book of Karlin and Lessard on sex ratio evolution (which, historically, was very important for the idea of long-term evolution). Karlin and Lessard realized that genetic constraints often prevent the evolution of an “evolutionarily stable” sex ratio (sensu Maynard Smith). Around this time, they collaborated with Eshel, and together the three considered long-term evolution. Karlin and Lessard decided to conduct a long-term simulation study by once in a while adding new genetic variants, in order to remove genetic constraints. This study resulted, however, in deception, as it did not converge at all. Sometimes the trajectory converged to a state close to the ESS sex ratio, but consecutively, it always moved away again. Later, I found the same in my own simulations, both in the context of sex ratio evolution and in the context of two-strategy games. Accordingly, I am much less convinced than Van Cleve that “long-term evolution” has been shown to “save” the ESS concept.

In contrast to the earlier sections, which are broadly accessible, Sections 6 and 7 are technically more sophisticated and will most likely only be understandable to experts. This holds, in particular, for Section 7. I wonder whether it is really useful to present Fig. 1 without discussing the underlying assumptions in sufficient depth and detail. The “reduction principle” is a wonderful tool but I would have liked to see a more critical discussion of its applicability.

Generally, the manuscript is well-written. But as mentioned above, there is a certain imbalance in this manuscript: no (restrictive) assumptions are stated in the first sections (although the experts would have liked to see them spelt out explicitly), while the treatment in the later sections is only understandable to experts in the field.

\end{reviewerquote}

\end{letter}

\end{document}